---
title: "Final code (final version)"
author: "Yineng Chen"
date: "2019/12/13"
output: html_document

---

These are codes we use in our final version:



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
library(tidyverse)
library(arsenal)
library(knitr)
require(survival)
library(patchwork)
```


# Data exploration: descriptive and visualization

```{r}
# read dataset
data_0=read_csv("Lawsuit.csv") %>% 
  janitor::clean_names()

# add variables: total salary and mean salary
data=data_0 %>% mutate(
  total_sal=sal94+sal95,
  mean_sal=total_sal/2
) %>% 
  # get rid of 94 salary, 95 salary and total salary
  dplyr::select(-sal94,-sal95,-total_sal)
  

# descriptive data

my_labels2 <- list(dept = "Deptartment,n%", clin="Area of emphasis n%", cert="Certification,n%",prate="Publication rate",exper="Years since obtaining MD",
                   mean_sal="Salary",rank="Rank,n%")

my_controls <- tableby.control(total = T,
  test=T,  
  digits=2,
  numeric.stats = c("meansd", "medianq1q3"),
  cat.stats = c("countpct"),
  stats.labels = list(
    meansd ="Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    countpct = "N (%)"))

data_1<-data %>% 
  mutate(dept=factor(dept, labels=c("Molecular Biology","Physiology","Genetics","Pediatrics","Medicine","Surgery")),
         gender=factor(gender, labels=c("Female","Male")),
         clin=factor(clin, labels=c("Primarily research emphasis","Primarily clinical emphasis")),
         cert=factor(cert, labels=c("not certified", "Board certified")),
         rank=factor(rank,label=c("Assistant","Associate","Full professor"))
         )

# Table 1
table <- tableby(gender ~ dept + clin+ cert + prate+exper+rank+mean_sal, data=data_1,control=my_controls)
c=summary(table, title = "Table1:Descriptive statistics for by Gender and Test", labelTranslations = my_labels2, text=T)
knitr::kable(c)
```



## distribution of y

```{r}
ori_y = ggplot(data_1,aes(x=mean_sal))+geom_histogram(aes(y=..density..)) + geom_density(col = "black",alpha=0)
```
not normal

```{r}
library(HH)

data_2=data_1 %>% 
  mutate(
    log_sal=log(mean_sal)
  )
ln_y = ggplot(data_2,aes(x=log_sal))+geom_histogram(aes(y=..density..)) + geom_density(col = "black",alpha=0)
data_2=data_2 %>% 
    dplyr::select(-id,-mean_sal)

ori_y + ln_y

```


# Visualization 
```{r}
library(patchwork)

rank_gen = data_1 %>%
  group_by(gender) %>%
  ggplot(aes(x = rank, y = mean_sal, color = gender)) + 
    geom_boxplot() 

dept_gen = data_1 %>%
  group_by(gender) %>%
  ggplot(aes(x = dept, y = mean_sal, color = gender)) + 
    geom_boxplot()

clin_gen = data_1 %>%
  group_by(gender) %>%
  ggplot(aes(x = clin, y = mean_sal, color = gender)) + 
    geom_boxplot()

cert_gen = data_1 %>%
  group_by(gender) %>%
  ggplot(aes(x = cert, y = mean_sal, color = gender)) + 
    geom_boxplot()

exp_gen = data_1 %>%
  group_by(gender) %>%
  ggplot(aes(x = exper, y = mean_sal, color = gender)) + 
    geom_point() +
  geom_smooth(method="lm",aes(color=gender),se=FALSE)

(rank_gen + dept_gen)/(clin_gen + cert_gen)

exp_gen


```


# confounder one by one

```{r}
gender_only=lm(log_sal~gender,data=data_2)
## coeffecient of gener is 0.3853

gender_dept=lm(log_sal~gender+dept,data=data_2)
summary(gender_dept)
#### coeffecient of gener changed from 0.3853 to  0.20521(>10%)
## dept is a confounder

gender_clin=lm(log_sal~gender+clin,data=data_2)
summary(gender_clin)
#### coeffecient of gener changed from 0.3853 to  0.33718(12.4%)
## clin is a confounder

gender_cert=lm(log_sal~gender+cert,data=data_2)
summary(gender_cert)
#### coeffecient of gener changed from 0.3853 to  0.33285(13.6%)
## cert is a confounder

gender_prate=lm(log_sal~gender+prate,data=data_2)
summary(gender_prate)
#### coeffecient of gener changed from 0.3853 to  0.251814(>10%)
## prate is a confouder

gender_exper=lm(log_sal~gender+exper,data=data_2)
summary(gender_exper)
#### coeffecient of gener changed from 0.3853 to  0.30755(>10%)
## exper is a confounder

gender_rank=lm(log_sal~gender+rank,data=data_2)
summary(gender_rank)
#### coeffecient of gener changed from 0.3853 to  0.349481(around 10%,9.2%)
## rank is  a potential confounder

mulit_2=lm(log_sal~gender+dept+clin+rank+cert+prate+exper,data=data_2)
summary(mulit_2)
## with very high adjR^2 in mulit_2

anova(gender_only,mulit_2)
## the p is smaller than 0.05
```



```{r}
## check the vif
vif(mulit_2)
## remove prate
##then the model is
gender_method1=lm(log_sal~gender+dept+clin+rank+cert+exper,data=data_2)
summary(gender_method1)
## but hard to find the stratified model, because the exper is continous variable
```

We are interested in the interaction effect of rank and experience.
 Is rank a modifier?
 Is experience a mpdifier?

#  Is rank a modifier? 
```{r}
gender_inter_rank=lm(log_sal~gender * rank + dept+clin+cert+exper,data=data_2)
summary(gender_inter_rank)


##  then stratified the rank levels
data_stra_assi=data_2 %>% 
  filter(rank=="Assistant")
data_stra_asso=data_2 %>% 
  filter(rank=="Associate")
data_stra_full=data_2 %>% 
  filter(rank=="Full professor")
## in assistant level
model_stra1=lm(log_sal~gender+dept+clin+cert+exper,data=data_stra_assi)
summary(model_stra1)

model_stra2=lm(log_sal~gender+dept+clin+cert+exper,data=data_stra_asso)
summary(model_stra2)

model_stra3=lm(log_sal~gender+dept+clin+cert+exper,data=data_stra_full)
summary(model_stra3)

# yes
```


#  Is experience a modifier? 
```{r}
gender_inter_exp = lm(log_sal~gender * exper + rank+dept+clin+cert+exper,data=data_2)
summary(gender_inter_exp )


gender_inter_two=lm(log_sal~gender * rank + dept+clin+cert+exper+gender*exper,data=data_2)
summary(gender_inter_two)
## after putting two interaction, the rank becomes non-sig, so don't put rank as an interaction
## but still put rank as counfounder.

```



## so our final model is
```{r}
gender_inter_exp=lm(log_sal~gender * exper +rank + dept + clin+cert,data=data_2)
summary(gender_inter_exp)
## stratify exp levels:
summary(data_2)
data_stra_exp=data_2 %>% 
  mutate(exper = as.factor(case_when(.$exper %in% c(0:5)~"very small", 
                            .$exper %in% c(6:10)~"small",
                            .$exper %in% c(11:14)~"median",
                            .$exper %in% c(15:37)~"large")))
gender_inter_exp_1=lm(log_sal~gender +rank + dept + clin+cert,data=data_stra_exp %>% filter(exper=="very small"))
summary(gender_inter_exp_1)
gender_inter_exp_2=lm(log_sal~gender +rank + dept + clin+cert,data=data_stra_exp %>% filter(exper=="small"))
summary(gender_inter_exp_2)
gender_inter_exp_3=lm(log_sal~gender +rank + dept + clin+cert,data=data_stra_exp %>% filter(exper=="median"))
summary(gender_inter_exp_3)
gender_inter_exp_4=lm(log_sal~gender +rank + dept + clin+cert,data=data_stra_exp %>% filter(exper=="large"))
summary(gender_inter_exp_4)
```


## diaganosis
```{r}
par(mfrow = c(2, 2))
plot(gender_inter_exp)
```
## meet the constant varience and normality assumption
## 184 maybe a potential influential point

## checking outlier and influential
### outlier in Y
```{r}
rstandard(gender_inter_exp)
```
Any observation with an absolute studentized residual ri greater than 2.5 is considered an outlier. In this dataset, we think 184 case is an outliers in Y.


### outlier in X
```{r}
hat=hatvalues(gender_inter_exp)
hat2=hat %>%
data.frame()
hat3=mutate(hat2,
hii=hat2[,1])
state_name=rownames(hat2) %>%
data.frame()
state_name2= mutate(state_name,
name=state_name[,1])
hat_final=cbind(cbind(dplyr::select(state_name2,name),dplyr::select(hat3,hii)))
hat_final
hat_final %>%
filter(hii>0.2)
```
## no outlier in X

## influential cases
```{r}
c=influence.measures(gender_inter_exp)
c
```

## try the model without outlier
```{r}
data_no_184=data_2[-184,]
gender_no_184=lm(log_sal~gender * exper +rank + dept + clin+cert,data=data_no_184)
summary(gender_no_184)
```
## model changed a lot so it is an influential point

```{r}
par(mfrow = c(2, 2))
plot(gender_no_184)
```


