---
title: "final"
author: "group 11 amazing"
date: "2019/12/3"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
```

## table1
```{r}
library(tidyverse)
library(arsenal)
library(knitr)
require(survival)
data_0=read_csv("Lawsuit.csv") %>% 
  janitor::clean_names()
data=data_0 %>% mutate(
  total_sal=sal94+sal95,
  mean_sal=total_sal/2
) %>% 
  select(-sal94,-sal95,-total_sal)
  
```

```{r}
my_labels2 <- list(dept = "Dept,n%", clin="Clin n%", cert="Cert,n%",prate="Publication rate",exper="Years since obtaining MD",
                   mean_sal="Salary",rank="Rank,n%")

 my_controls <- tableby.control(total = T,
  test=T,  
  digits=4,
  numeric.stats = c("meansd", "medianq1q3"),
  cat.stats = c("countpct"),
  stats.labels = list(
    meansd ="Mean (SD)",
    medianq1q3 = "Median (Q1, Q3)",
    countpct = "N (%)"))
```

```{r}
data_1<-data %>% 
  mutate(dept=factor(dept, labels=c("Molecular Biology","Physiology","Genetics","Pediatrics","Medicine","Surgery")),
         gender=factor(gender, labels=c("Male","Female")),
         clin=factor(clin, labels=c("Primarily clinical emphasis,","Primarily research emphasis")),
         cert=factor(cert, labels=c("Board certified","not certified")),
         rank=factor(rank,label=c("Assistant","Associate","Full professor"))
         )
```


```{r}

table <- tableby(gender ~ dept + clin+ cert + prate+exper+rank+mean_sal, data=data_1,control=my_controls)
c=summary(table, title = "Table1:Descriptive statistics for by Gender and Test", labelTranslations = my_labels2, text=T)
knitr::kable(c)
```

## distribution of y

```{r}
ggplot(data_1,aes(x=mean_sal))+geom_histogram(aes(y=..density..)) + geom_density(col = "black",alpha=0)
```
not normal

```{r}
library(HH)
data_2=data_1 %>% 
  mutate(
    log_sal=log(mean_sal)
  )
  ggplot(data_2,aes(x=log_sal))+geom_histogram(aes(y=..density..)) + geom_density(col = "black",alpha=0)
data_2=data_2 %>% 
    dplyr::select(-id,-mean_sal)
```

seems normal

```{r}
# stepwise can not be used in association model, only in predictive model
muli1=lm(log_sal~.,data=data_2)
step(muli1) %>% 
  summary()
```

```{r}
muli2=lm(formula = log_sal ~ gender + dept + clin + cert + prate + exper + rank, 
    data = data_2)
summary(muli2)
muli3=lm(formula = log_sal ~ gender + dept + clin + cert + exper + rank, 
    data = data_2)
summary(muli3)
(0.025763-0.019536)/0.019536
```

```{r}
# foreward
full_model = lm(log_sal ~ . ,data = data_2)
step_null <- lm(log_sal ~1, data = data_2)
for_model <- step(step_null, scope = list(lower = step_null, upper = full_model), direction = "forward")
```


```{r}
# colinearity
library(rms)
vif(muli1)
## build the model then remove correlated prate
```


# correlation
pairs(data_2)
# normality
library(DataExplorer)
plot_qq(data_2)
# categorical
plot_bar(data_2, maxcat = 20, parallel = TRUE)


#method 1, try the interaction along with confounders
## try to find confounder
## put all variable into model-dept, clin,cert, prate,exper,rank

```{r}
gender_only=lm(log_sal~gender,data=data_2)
## coeffecient of gener is 0.3853

gender_dept=lm(log_sal~gender+dept,data=data_2)
summary(gender_dept)
#### coeffecient of gener changed from 0.3853 to  0.20521(>10%)
## dept is a confounder

gender_clin=lm(log_sal~gender+clin,data=data_2)
summary(gender_clin)
#### coeffecient of gener changed from 0.3853 to  0.33718(12.4%)
## clin is a confounder

gender_cert=lm(log_sal~gender+cert,data=data_2)
summary(gender_cert)
#### coeffecient of gener changed from 0.3853 to  0.33285(13.6%)
## cert is a confounder

gender_prate=lm(log_sal~gender+prate,data=data_2)
summary(gender_prate)
#### coeffecient of gener changed from 0.3853 to  0.251814(>10%)
## prate is a confouder

gender_exper=lm(log_sal~gender+exper,data=data_2)
summary(gender_exper)
#### coeffecient of gener changed from 0.3853 to  0.30755(>10%)
## exper is a confounder

gender_rank=lm(log_sal~gender+rank,data=data_2)
summary(gender_rank)
#### coeffecient of gener changed from 0.3853 to  0.349481(<10%)
## rank is not a confounder

mulit_2=lm(log_sal~gender+dept+clin+cert+prate+exper,data=data_2)
summary(mulit_2)
## with very high adjR^2 in mulit_2

anova(gender_only,mulit_2)
## the p is smaller than 0.05
```


## try the interaction 
```{r}
gender_inter_dept=lm(log_sal~gender*dept+clin+cert+prate+exper,data=data_2)
summary(gender_inter_dept)
## the interacton term are not sig

gender_inter_clin=lm(log_sal~gender*clin+dept+cert+prate+exper,data=data_2)
summary(gender_inter_clin)
## the interacton term is sig

gender_inter_cert=lm(log_sal~gender*cert+clin+dept+prate+exper,data=data_2)
summary(gender_inter_cert)
## the interacton term isn't sig

gender_inter_prate=lm(log_sal~gender*prate+clin+dept+cert+exper,data=data_2)
summary(gender_inter_prate)
## the interacton term isn't sig

gender_inter_exper=lm(log_sal~gender*exper+clin+dept+cert+prate,data=data_2)
summary(gender_inter_exper)
## the interacton term is sig


gender_inter=lm(log_sal~gender*clin+gender*exper+dept+cert+prate,data=data_2)
summary(gender_inter)

## check the vif
vif(gender_inter)
## remove prate
##then the model is
gender_method1=lm(log_sal~gender*clin+gender*exper+dept+cert,data=data_2)
summary(gender_method1)
## but hard to find the stratified model, because the exper is continous variable
```

# methos 2 try confounders in above same way, but find the interaction independtly.
```{r}
gender_inter1=lm(log_sal~gender*dept,data=data_2)
summary(gender_inter1)
## interaction term  non-sig
gender_inter2=lm(log_sal~gender*clin,data=data_2)
summary(gender_inter2)
## interaction term  non-sig
gender_inter3=lm(log_sal~gender*cert,data=data_2)
summary(gender_inter3)
## interaction term  non-sig
gender_inter4=lm(log_sal~gender*prate,data=data_2)
summary(gender_inter4)
## interaction term  non-sig
gender_inter5=lm(log_sal~gender*exper,data=data_2)
summary(gender_inter5)
## interaction term  non-sig
gender_inter6=lm(log_sal~gender*rank,data=data_2)
summary(gender_inter6)
## interaction term sig

gender_method2=lm(log_sal~gender*rank+dept+clin+cert+exper,data=data_2)
summary(gender_method2)


## the stratified models are due to the rank levels
data_stra_assi=data_2 %>% 
  filter(rank=="Assistant")
data_stra_asso=data_2 %>% 
  filter(rank=="Associate")
data_stra_full=data_2 %>% 
  filter(rank=="Full professor")
## in assistant level
model_stra1=lm(log_sal~gender+dept+clin+cert+exper,data=data_stra_assi)
summary(model_stra1)

model_stra2=lm(log_sal~gender+dept+clin+cert+exper,data=data_stra_asso)
summary(model_stra2)

model_stra3=lm(log_sal~gender+dept+clin+cert+exper,data=data_stra_full)
summary(model_stra3)
```


